---
title: "Clase 2"
author: "Erick Cuevas Fernández"
date: "25/8/2019"
output: pdf_document
---
# Documentos csv
**comma separate values**

*read.csv*

Funciones que leen .csv 

- `read.csv(file_path)`
- `read_csv(file_path)`

**Argumentos de la función**

`read.csv(file, header = TRUE, sep = ",", quote = "\"", dec = ".", fill = TRUE, comment.char = "", ...)`

`read.csv2(file, header = TRUE, sep = ";", quote = "\"", dec = ",", fill = TRUE, comment.char = "", ...)`

**Tambien podemos leer archivos separados por espacio y tabuladores:**

`read.table(file, header = FALSE, sep = "", quote = "\"'", dec = ".", numerals = c("allow.loss", "warn.loss", "no.loss"), row.names, col.names, as.is = !stringsAsFactors, na.strings = "NA", colClasses = NA, nrows = -1, skip = 0, check.names = TRUE, fill = !blank.lines.skip, strip.white = FALSE, blank.lines.skip = TRUE, comment.char = "#", allowEscapes = FALSE, flush = FALSE, stringsAsFactors = default.stringsAsFactors(), fileEncoding = "", encoding = "unknown", text, skipNul = FALSE)`

`read.delim(file, header = TRUE, sep = "\t", quote = "\"", dec = ".", fill = TRUE, comment.char = "", ...)`

`read.delim2(file, header = TRUE, sep = "\t", quote = "\"", dec = ",", fill = TRUE, comment.char = "", ...)`

# Ejemplo para leer un archivo .csv
```{r}
names_data <- c("age", "sex", "cp", "trestbps", "chol", "fbs", "restecg",
                "thalach", "exang", "oldpeak", "slope", "ca", "thal", "num")
data_cleveland <- read.csv("Datasets/processed.cleveland.data")
colnames(data_cleveland) <- names_data

str(data_cleveland)

```

Attribute Information:

- age: age in years

- sex: sex (1 = male; 0 = female)

- cp: chest pain type

        -- Value 1: typical angina
        -- Value 2: atypical angina
        -- Value 3: non-anginal pain
        -- Value 4: asymptomatic
        
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)

- chol: serum cholestoral in mg/dl

- fbs: (fasting blood sugar > 120 mg/dl)  (1 = true; 0 = false)

- restecg: resting electrocardiographic results

        -- Value 0: normal
        -- Value 1: having ST-T wave abnormality (T wave inversions and/or ST 
                    elevation or depression of > 0.05 mV)
        -- Value 2: showing probable or definite left ventricular hypertrophy
                    by Estes' criteria

- thalach: maximum heart rate achieved

- exang: exercise induced angina (1 = yes; 0 = no)

- oldpeak = ST depression induced by exercise relative to rest

- slope: the slope of the peak exercise ST segment

        -- Value 1: upsloping
        -- Value 2: flat
        -- Value 3: downsloping
        
- ca: number of major vessels (0-3) colored by flourosopy

- thal: 3 = normal; 6 = fixed defect; 7 = reversable defect

- num: diagnosis of heart disease (angiographic disease status)

        -- Value 0: < 50% diameter narrowing
        -- Value 1: > 50% diameter narrowing

*Source Information:*

   (a) Creators: 
   
       -- 1. Hungarian Institute of Cardiology. Budapest: Andras Janosi, M.D.
       
       -- 2. University Hospital, Zurich, Switzerland: William Steinbrunn, M.D.
       
       -- 3. University Hospital, Basel, Switzerland: Matthias Pfisterer, M.D.
       
       -- 4. V.A. Medical Center, Long Beach and Cleveland Clinic Foundation:
             Robert Detrano, M.D., Ph.D.
             
   (b) Donor: David W. Aha (aha@ics.uci.edu) (714) 856-8779   
   
   (c) Date: July, 1988*

# Documentos json

**.json**

**JavaScript Object Notation**

El tipo de archivo con mayor uso en las bases de datos, ya que consumo muy poco espacio de memoria. La paquetería *jsonlite* nos permite leer este tipo de archivos en R.

```{r}
library(jsonlite)

```

*fromJSON()*

```{r}
id <- 681  
url_chem <- paste0("https://pubchem.ncbi.nlm.nih.gov/rest/pug_view/data/compound/"
                   ,id,
                   "/json")

response <- fromJSON(url_chem)

# str(response)

archivo_sections<-response$Record$Section$Section

```

Desde JSON obtenemos una lista o un data.frame, y ya que sabemos manejar listas, podemos empezar a obtener los datos que deseamos.



# Documentos xml

# Documentos txt
```{r}
texto <- readLines("Datasets/ejemplo.txt")
head(texto)

```


# Documentos SPSS
